Transforming Trash as an Artistic Act

### 2016-01-17 chapter1_final.tex
turnitin fixes
Former-commit-id: 18946952758077b82f1e3365219cc043e9cc2f0d

% Different scales
The scale of trash production is not same for everybody or every country. \citet[2]{chen2015waste} writes that by referring to analysis of World Bank \quotes{the amount of solid waste generated in cities is growing faster than the rate of urbanization.} She argues that when the people's income level and the rate of urbanization increased, production of waste also increases. Chen notes that nearly half of the world’s waste are produced by OECD\footnote{OECD (Organization for Economic Co-operation and Development) is an international economic organization mostly composed of the western developed countries. Turkey is a member of this organization.} countries. On the other hand, Africa and South Asia generate least waste. Similar to these as stated by \citet[16]{cerny1996recycled} \quotes{a person’s wealth has become measured not only in how much he or she can afford to consume, but in how much he or she can afford to throw away.} Moreover from trash many unexpected results can be extracted because as \citet[xxv]{zimring2012encyclopedia} argues that:
\begin{quote}
Our trash is a testament; what we throw away says much about our values, our habits, and our lives. ... Our trash is part of us, whether or not we choose to acknowledge it. ... The absence of a waste stream aroused suspicion, just as the presence of particular items tell us about the habits of the consumers who generate a waste stream.
\end{quote}

@@

According to dictionary Merriam-Webster, the word reuse means \quotes{to use again especially in a different way or after reclaiming or reprocessing.} Reusing includes usage of the same product unaltered in structure. Reusing lengthens the life of the item or material. To reuse is to use something again instead of throwing it away. In this thesis, reuse is explored through the work of the artist.

On the other hand recycle means \quotes{to treat or process (used or waste materials) so as to make suitable for reuse.} Particularly, through the process of recycling, an object is used as a raw material in the production of new product. Recycling occurs when waste in an unchanged chemical form is used in the same process that created the original product. Crushed glass containers used to make new glass containers, and scrap metal used in foundries can be given as examples.

@@

\citet[72]{strasser1999waste} explains recycling by referring to the Oxford English Dictionary as:
\begin{quote}
... recycling originated in the oil industry during the 1920s, describing partially refined petroleum sent through the refining cycle again to reduce waste. Its use broadened during the 1960s, designating all kinds of reuse and reclamation, and the world became familiar during the early 1970s, as the burgeoning environmental movement promoted the separate collection of certain kinds of trash to promote their reuse in manufacturing.
\end{quote}

Some of the scholars use recycling \citep{cerny1996recycled,herman1998trashformations} for the recreation and transformation of waste. Many scholar used the word recycling when mentioning works use trash. However they are not mentioning the meaning that is decomposing things to the their particles. What they actually referring is reusing and combining things, concepts to create new mixtures. It is hard to say that they use it wrongly, but what they refer is actually upcycling. Creative reuse, inventing new things from discard. Upcycling is a term coined by architect and designer William McDonaugh and chemist Michael Braungart and refers to \quotes{the process of converting an industrial nutrient (material) into something of similar or greater value, in its second life} \citep[as cited in][63]{emgin2012trashion}. It is argued that art, in this scope, acts as an instrument of transformation and re-presents artifacts what was once regarded waste. This approach resists the argument that an object is dead once it is disposed of \citep{emgin2012trashion}. Some of the scholars and artist uses the method of upcycling for their works. Undertaking thesis project is about those objects that are reproduced from trash through the procedure of upcycling.

>>>>

% Different scales
The scale of trash production is not same for everybody or every country. \citet[2]{chen2015waste} writes that by referring to analysis of World Bank \quotes{the amount of solid waste generated in cities is growing faster than the rate of urbanization.} She argues that when the people's income level and the rate of urbanization increased, production of waste also increases. Chen notes that nearly half of the world’s waste are produced by OECD\footnote{OECD (Organization for Economic Co-operation and Development) is an international economic organization whose members are mostly western developed countries.} countries. On the other hand, Africa and South Asia generate least waste. Similar to these as stated by \citet[16]{cerny1996recycled} \quotes{a person’s wealth has become measured not only in how much he or she can afford to consume, but in how much he or she can afford to throw away.} Moreover from trash many unexpected results can be extracted because as \citet[xxv]{zimring2012encyclopedia} argues that:

\begin{quote}
Our trash is a testament; what we throw away says much about our values, our habits, and our lives. ... Our trash is part of us, whether or not we choose to acknowledge it. ... The absence of a waste stream aroused suspicion, just as the presence of particular items tell us about the habits of the consumers who generate a waste stream.
\end{quote}

@@

According to dictionary Merriam-Webster, the word reuse means \quotes{to use again especially in a different way or after reclaiming or reprocessing.} Reusing includes usage of the same product unaltered in structure. Reusing lengthens the life of the item or material in society. To reuse is to use something again instead of throwing it away. In this thesis, reuse is explored through the work of the artist.

On the other hand recycle means \quotes{to treat or process (used or waste materials) so as to make suitable for reuse.} Particularly, through the systematic process of recycling, an object is used as a raw material in the production of new product. Recycling occurs when waste in an unchanged chemical form is used in the same process that created the original product. Crushed glass containers used to make new glass containers, and scrap metal used in foundries can be given as examples.

@@

\citet[72]{strasser1999waste} by referring to the Oxford English Dictionary notes that \quotes{recycling originated in the oil industry during the 1920s.} Here the main purpose is to reduce the waste of petroleum therefore \quotes{partially refined petroleum sent through the refining cycle again.} She explains that \quotes{the world became familiar during the early 1970s, as the burgeoning environmental movement promoted the separate collection of certain kinds of trash to promote their reuse in manufacturing.}

Some of the scholars use recycling \citep{cerny1996recycled,herman1998trashformations} for the recreation and transformation of waste. Many scholar used the word recycling when mentioning works use trash. However they are not mentioning the meaning that is decomposing things to the their particles. What they actually referring is reusing and combining things, concepts to create new mixtures. It is hard to say that they use it wrongly, but what they refer is actually upcycling. Creative reuse, inventing new things from discard. The term upcycling is coined by architect and designer William McDonaugh and chemist Michael Braungart. It can be explained as \quotes{the process of converting an industrial nutrient (material) into something of similar or greater value, in its second life} \citep[as cited in][63]{emgin2012trashion}.  Upcycling approach resists the argument that an object is dead once it is disposed of \citep{emgin2012trashion}. It can be argued that art, in this scope, acts as an instrument of transformation and re-presents artifacts what was once regarded waste. In the light of this argument, methods of analyzed artists in this thesis when producing their artworks can be categorized as upcycling. Additionally, undertaking thesis project is about handmade notebooks that are reproduced from discarded papers through the procedure of upcycling.

### 2016-01-17 chapter2_final.tex
turnitin fixes
Former-commit-id: 18946952758077b82f1e3365219cc043e9cc2f0d

@@

Throw away culture has a significant place in producing trash. It is a behavioral pattern, and an approach to objects and commodities. Some artists and people develop different approaches to them. In this type of culture or behavior throwing them is a natural act and promoted. Consumption and wasting things are encouraged. On the other hand, there are examples of saving everything instead of wasting. In the work of Song Dong, \textit{Waste Not} thousands of domestic objects are exhibited. They are owned by his mother and saved for later use. They are extracted from her home. The compiled objects ranging from toothpaste to shoe, from bottles to toys. His mother lived a great depression in China, and suffered from poorness. As a result of it she developed the habit of saving things. After died of her husband this habit reached to extreme levels. She started to save every tiny object. For her, trash becomes an essential part of her life. In fact, it not is hard to say that preserved these things are not useless or worthless. She establishes a different relationship with these used objects.

@@

Jane Alison, senior curator at the Barbican, states that \textit{Waste Not} \quotes{helps us to understand the reality of Chinese history and culture in the twentieth century century in a way that newspapers can’t} \citep[as cited in][]{needham2015chinese}

As it can be understood trash reflects people's habits. It is like a mirror. It can not be separated from their producers and it is well studied area by scholars. Garbology is a study of waste as a social science. It applies methodologies of archeology to the human debris. Weberman infamously used techniques of what he deemed garbology to uncover what he saw as the essential nature of people. He once said, perhaps indirectly referencing Jean Brillat-Savarin’s quote about food, \quotes{You are what you throwaway} \citep[as cited in][301]{lukas2012garbage}. The field of garbology involves the study of refuse and waste. It enables researchers to document information on the dump and changing patterns of modern rubbish, hence assisting in the study of contemporary human society or culture \citep{roy2012garbology}. According to the Oxford English Dictionary, the term was first used by waste collectors in the 1960s. Weberman popularized the term in describing his study of Bob Dylan’s garbage in 1970. It was pioneered as an academic discipline by William Rathje at the University of Arizona in 1973. The work of archaeologists such as William Rathje and Cullen Murphy has offered significant insights into the relationship between archaeological artifacts and garbage, exploring the function of trash as a resource for understanding the cultural and social practice. At the same time, thinkers such as Zygmunt Bauman and Giorgio Agamben have shed light on the fate of the human being as a wasted or discarded element in discourses of socio-political hygiene \citep{pye2010trashculture}. With the help of mentioned understanding of trash, what can be understand from the works of art?

@@

Tracey Emin puts her private life to the public. This work can be as a portrait of artist in a post-modern fashion. Around the bed, there are pieces of junks speared all over the room. It can be understood from her junk what type of life she has. To put it another way, viewer enter the world of the artist through her trashes.

% From wiki
Every detail of her private life is presented through her bed in an uncensored fashion. The idea of \textit{My Bed} was result of a depressive period in the artist’s life when she had remained in bed for several days without eating or drinking anything but alcohol. The sensation of depression leading to disarray and detritus is not completely unusual to many people, and \textit{My Bed} therefore enforces the communicative power that Tracey Emin seeks with her work. The artwork generated great media enthusiasm, particularly over the fact that the bed sheets were stained with bodily secretions and the floor had items from the artist’s room, such as condoms, underwear with menstrual blood stains, other junks, and functional, everyday objects, including a pair of slippers. The bed was presented in the state that Emin claimed it had been after languishing in it for several days; at the time she was suffering suicidal depression brought on by relationship difficulties.

It is a realization what she did during her depression time. It perfectly represents her condition during those days. It can be understood that trash is part of our life. Things that are exhibited in this work are part of her action. She created them and later through them in the mind of people the condition and life of Tracey Emin are recreated.

The work can be interpreted as a self-portrait of a woman at a time of emotional depression. However, it is still a bed and all that a bed symbolizes and covers: sleep, sleeplessness, sex, birth, death and dreams. Particularly for the ones that have small rooms and spaces bed becomes a life space. I have also experienced such a thing at boarding school and university dormitory for total my eight years. Having tight space sometimes bed becomes a place for studying and reading. Sometimes it was used to dry the clothes. It is more than sleeping and provide place for other activities.

% tracey emin joseph beuys arasında bir ilişki var. sanatsal eylem ve kalıntı olarak bakıldığında.

% Aslan yattigi yerden belli olur.

@@

%****************************************
\section{Rubbish Theory}
Objects have a lifetime similar to humans, and they do not remain same through that time. They move around, and their value, usage, location change constantly. During their lifetime objects may circulate different markets and values systems such as economic, social and aesthetic. Especially these cycles have accelerated with the arrival of consumer culture. For instance, most recent technological gadgets become obsolete within two or three years. Firsthand, secondhand and even trash is exchanged among not only people but also countries. People from other cultures and understanding interpret objects differently. They often are not aware of their initial and intended usage. They find new original purposes and meanings. Therefore, objects turn different items in their hand. Anthropologist Sidney Mintz \citep[as cited in][]{strasser1999waste} writes that \quotes{I have seen automobile bearings fashioned into portable vulcanizing kits; bits of toothbrush handles used as \singlequotes{jewels} for rings, and ordinary tin cans turned into simple kerosene lamps.} As it can be understood that objects function and value are transformed by relocation and revaluation of objects from one place to the other or one discipline to another.

It can be claimed that the end of an object’s life equates to the moment when it is refused. This action might take place in different forms and for various reasons. Nonetheless, in the most accurate and common sense, the life of an object ends in a dump in the state of waste. At this moment, the object is left valueless in all the possible meanings of the term value. It can no more serve a function, it can on no account be exchanged for anything else, and it can by no means engage in the processes of signification to connote and endow its user with specific social values \citep[63]{emgin2012trashion}. However, Thompson’s Rubbish Theory shows us the opposite of it. Trash can is not an end it is starting point of another life.

>>>>

%At the same time, thinkers such as Zygmunt Bauman and Giorgio Agamben have shed light on the fate of the human being as a wasted or discarded element in discourses of socio-political hygiene \citep{pye2010trashculture}.

@@

Throw away culture has a significant place in producing trash. It is a behavioral pattern, and an approach to objects and commodities. Some artists and people develop different approaches to them. In this type of culture or behavior throwing them is a natural act and promoted. Consumption and wasting things are encouraged. On the other hand, there are examples of saving everything instead of wasting. In the work of Song Dong, \textit{Waste Not}, thousands of domestic objects are exhibited. They are owned by his mother and saved for later use. They are extracted from her home. The compiled objects ranging from toothpaste to shoe, from bottles to toys. His mother lived a great depression in China, and suffered from poorness. As a result of it she developed the habit of saving things. After died of her husband this habit reached to extreme levels. She started to save every tiny object. For her, trash becomes an essential part of her life. In fact, it not is hard to say that preserved these things are not useless or worthless. She establishes a different relationship with these used objects.

@@

Jane Alison, senior curator at the Barbican Art Gallery in London, states that \quotes{\textit{Waste Not} helps us to understand the reality of Chinese history and culture in the twentieth century century in a way that newspapers can’t} \citep[as cited in][]{needham2015chinese}

As it can be understood trash reflects people's habits. It is like a mirror. It can not be separated from their producers and it is well studied area by scholars. Garbology is a study of waste as a social science. It applies methodologies of archeology to the human debris. Weberman infamously used techniques of what he deemed garbology to uncover what he saw as the essential nature of people. He once said, perhaps indirectly referencing Jean Brillat-Savarin’s quote about food, \quotes{You are what you throwaway} \citep[as cited in][301]{lukas2012garbage}. The field of garbology involves the study of refuse and waste. It enables researchers to document information on the dump and changing patterns of modern rubbish, hence assisting in the study of contemporary human society or culture \citep{roy2012garbology}. According to the Oxford English Dictionary, the term was first used by waste collectors in the 1960s. Weberman popularized the term in describing his study of Bob Dylan’s garbage in 1970. It was pioneered as an academic discipline by William Rathje at the University of Arizona in 1973. The work of archaeologists such as William Rathje and Cullen Murphy has offered significant insights into the relationship between archaeological artifacts and garbage, exploring the function of trash as a resource for understanding the cultural and social practice. With the help of mentioned understanding of trash, what can be understand from the works of art?

@@

Tracey Emin puts her private life to the public via representation of her bed. This work can be as a portrait of artist in a post-modern fashion. Around the bed, there are pieces of junks speared all over the room. It can be understood from her junk what type of life she has. To put it another way, viewer enter the world of the artist through her trashes.

Every detail of Emin's private life is presented through her bed in an uncensored fashion. The idea of \textit{My Bed} (Figure \ref{fig:TraceyEmin_MyBed}) was result of a depressive period in the artist’s life when she had stayed in bed for a few days without eating or drinking anything except alcohol. The feeling of depression resulting in mess and detritus is not completely unusual to many people, and \textit{My Bed} therefore has a bold affect on the viewer. The artwork generated great media enthusiasm, particularly over the fact that the bed sheets were stained with bodily secretions and the floor had items from the artist’s room, such as condoms, underwear with menstrual blood stains, other junks, and functional, everyday objects, including a pair of slippers.

It is a realization what she did during her depression time and perfectly represents her condition during those days. It can be understood that trash is part of our life. Things that are exhibited in this work are part of her action. She created them and later through them in the mind of people the condition and life of Tracey Emin are recreated.

The work can be interpreted as a self-portrait of a woman at a time of emotional depression. However, it is still a bed and all that a bed symbolizes and covers life events such as sleep, sleeplessness, sex, birth, death and dreams. Further bedroom is one of the most personal spaces in the homes. Particularly for the ones that have small rooms and spaces bed becomes a life space. I have also experienced such a thing at boarding school and university dormitory for total my eight years. Having tight space sometimes bed becomes a place for studying and reading. Sometimes it was used to dry the clothes. It is more than sleeping and provide place for other activities.

@@

%****************************************
\section{Rubbish Theory}
Objects have a lifetime similar to humans, and they do not remain same through that time. They move around, and their value, usage, location change constantly. During their lifetime objects may circulate different markets and values systems such as economic, social and aesthetic. Especially these cycles have accelerated with the arrival of consumer culture. For instance, most recent technological devices become outdated within a few years. Firsthand, secondhand and even trash is exchanged among not only people but also countries. People from other cultures and understanding interpret objects differently. They often are not aware of their initial and intended usage. They find new original purposes and meanings. Therefore, objects turn different items in their hand. Anthropologist Sidney Mintz \citep[as cited in][]{strasser1999waste} writes that \quotes{I have seen automobile bearings fashioned into portable vulcanizing kits; bits of toothbrush handles used as \singlequotes{jewels} for rings, and ordinary tin cans turned into simple kerosene lamps.} As it can be understood that objects function and value are transformed by relocation and revaluation of objects from one place to the other or one discipline to another.

It can be claimed that the end of an object’s life equates to the moment when it is refused. As stated by \citet[63]{emgin2012trashion} this action \quotes{might take place in different forms and for [various] reasons.} In common sense it is considered as waste. At this moment, the object is left valueless in all the possible meanings of the term value. It can no more serve a function, it can on no account be exchanged for anything else, and it can by no means engage in the processes of signification to connote and endow its user with specific social values \citep[63]{emgin2012trashion}. However, Thompson’s Rubbish Theory shows us the opposite of it. Trash can is not an end it is starting point of another life.
